---
title: "Format of input dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_format}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`SLCARE` requires the input datset have a specific format. We will illustrate the format with the `SLCARE_simdat` from the `SLCARE` package.

```{r}
library(SLCARE)
packageVersion("SLCARE")
```

```{r}
data("SLCARE_simdat", package = "SLCARE")
head(SLCARE_simdat)
```

Dataset `SLCARE_simdat` consists of 5 columns. We can devide the 5 columns into 4 parts:

- Column 1 `ID`: Subjects identification
- Column 2 `time`: Time for recurrent event and censoring time (which usually represent the longest follow up time)
- Column 3 `event`: Recurrent event indicator. A value of `1` represents the recurrent event is recorded; a value of `0` represents censoring.
- Columns 4-5 `x1` and `x2`: time independent corvariates.

Input datset should have the same format as `SLCARE_simdat`. Specifically, column 1-3 must be `ID`, `time` and `event`, and they must in the same order. Input datset can have as many corvariates as you need. They should be placed from column 4 which locates after `ID`, `time` and `event`.


