---
title: "SLCARE"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SLCARE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`SLCARE` models recurrent events data based on a semi-parametric multiplicative model. 

## Data and Model assumption

Let $T_{i} ^{(j)}$ denote time to the $j^{th}$ recurrent event, $\tilde{Z} _{i}$ denote a $p \times 1$ vector of time-independent covariates, $C_{i}$ censoring time.

Counting process for recurrent events:
$$
N_{i} ^{*} (t) = \sum _{j = 1} ^{\infty} I(T_{i} ^{(j)} \leq t), i = 1,\cdots ,n
$$
Assume $N_{i} ^{*} (t)$ is a non-stationary Poisson process with intensity function:
$$
\lambda _{i} (t) = \sum _{k = 1} ^{K} I (\xi _{i} = k) \times \lambda _{0} (t) \times W_{i} \times \eta _{0,k} \times exp(\tilde{Z} _{i} ^{T} \tilde{\beta} _{0,k})\tag{1}
$$
Where:

$K$ : a pre-determined number of latent class

$\xi _{i}$: un-observable latent class membership

$\lambda _{0} (t)$:  an unspecified, continuous, non-negative baseline intensity function shared among different latent class

$W_{i}$: a positive subject-specific latent variable independent of $(\xi, \tilde{Z}, C)$, $W_{i}$ offers the flexibility to accommodate individual difference with a larger or smaller value indicating more or less frequent occurence of recurrent events.

Unknown class-specific parameters:

$\eta _{0,k}$ Captures the class-k scale shift in the baseline intensity function

$\tilde{\beta} _{0,k}$ represents the class-k corvariate effects on the intensity function

To ensure the identifiability of $\lambda _{0} (t)$ And $\eta _{0,k}$ in (1), assume $E(W_{i} | \tilde{Z}_{i} , \xi _{i} = k) = 1$ for $k = 1, \cdots, K$ and impose the constraint:
$$
\int _{0} ^{v^{*}} \lambda _{0} (t) dt = 1 \tag{2}
$$
where $v^{*}$ is a predetermined constant

- In practice, it may be chosen to be slightly smaller than the upper bound of $C_{i}$ 's support
- Or it would only imply a scale shift to $\lambda _0 (t)$ by a constant with $\eta _{0,k}$ and $\tilde{\beta} _{0,k}$ remaining the same

To address the difficulty with the unobservable latent class membership, assume a multinomial logistic regression model for $\xi _{i}$ :
$$
P(\xi _{i} = k | \tilde{Z} _{i}) = p_{k} (\alpha _{0} , \tilde{Z} _{i}) = \frac{exp(\tilde{Z} _{i} ^{T} \alpha _{0,k})}{\sum_{k = 1}^{K}exp(\tilde{Z} _{i} ^{T} \alpha _{0,k}) } , k = 1, \cdots, K \tag{3}
$$
Where: $\alpha _0 = (\alpha _{0,1} ^{K}, \dots, \alpha _{0,K}^{K})^{T}$ 

## Examples
We will illustrate the usage of `SLCARE` with `SLCARE_simdat` from the `SLCARE` package.

```{r}
library(SLCARE)
data("SLCARE_simdat", package = "SLCARE")
head(SLCARE_simdat)
```
### Fit the model with K = 2 with frailty $W = 1$. 
Note that by default, the initial values in the iteration based estimation procedure are derived with K- means and a multiplicative intensity model studies by Wang et al. (2001) using the reReg() function. See `SLCARE::get_initial()` for details.
```{r}
model1 <- SLCARE(dat = SLCARE_simdat, K=2)
```

`SLCARE` will return to a list containing:
```{r}
names(model1)
```

We can check the point estimates for $\alpha$ and $\beta$
```{r}
model1$alpha
model1$beta
```

We can check the posterior prediction for observed events for subjects of interest
```{r}
model1$PosteriorPrediction
```

We can check the posterior probability of latent class membership
```{r}
model1$EstimatedTau
```

We can check the plot for model checking
```{r}
model1$ModelChecking
```

We can check the plot for $\hat{\mu} _{0}(t)$
```{r}
model1$Estimated_mu0t
```

We can check the $\hat{\mu} _{0} (t)$ at specific time points. For example, let $t = 100, 1000, 5000$
```{r}
model1$est_mu0(c(100, 1000, 5000))
```

We can check the plot of estimated mean function
```{r}
model1$Estimated_Mean_Function
```

We can check the relative entropy for the fitted model
```{r}
model1$RelativeEntropy
```

We can check the initial values for iteration based estimation procedure
```{r}
model1$InitialAlpha
model1$InitialBeta
```

### Fit the model with K = 2 with frailty $W = 1$ and apply self-defined initial values in estimation procedure.

```{r}
alpha <- matrix(c(0, 0, 0.5, -2, 2, -4), nrow = 3, ncol = 2, byrow = TRUE)
beta <- matrix(c(2.5, -0.5, -0.3, 1.5, -0.2, -0.5, 2.5,  0.1, 0.2), nrow = 3 , ncol = 2+1 , byrow = TRUE)
model2 <- SLCARE(alpha, beta, dat = SLCARE_simdat)
```

### Fit the model with K = 2 with frailty $W \sim gamma(3,3)$ and apply self-defined initial values in estimation procedure.

```{r}
model3 <- SLCARE(alpha, beta, dat = SLCARE_simdat, gamma = 3)
```

We may also calculate bootstrap standard errors by using `boot` argument
```{r}
# model4 <- SLCARE(alpha, beta, dat = SLCARE_simdat, boot = 100)
```

We can check bootstrap standard errors
```{r}
# model4$alpha_bootse
# model4$beta_bootse
```
